---
title: "Checkpoint3"
author: "Ruben Jimenez"
format: html
editor: source
embed-resources: true
code-tools: true
---

The measles virus affects hundreds of thousands or even millions annually despite the global distribution of an effective vaccine. We are interested in exploring the socioeconomic and geographic factors that may influence infection rates.

# Load packages

```{r setup}
#| include: false
#| message: false
library(httr)
library(tidyjson)
library(jsonlite)
library(tidyverse)
library(glue)
library(tidytuesdayR)
library(gt)
library(rvest)

measles_year <- read_csv("measles_year.csv") |>
  select(-1)

```

# Pull data from Restful Countries API

```{r}

world_df <- GET("https://restcountries.com/v4/all?fields=cca3,geolocation,area,borders,hdi,gini,government,religion") |> (\(x) rawToChar(x$content))() |> fromJSON() |>
  unnest_wider(c(gini, geolocation, government, religion), "_") |>
  relocate(cca3) |>
  rename("ISO3" = cca3) |>
  arrange(ISO3)

world_df |>
  head(5) |>
  gt() |> tab_header(title = md("REST API Data")) |>
    cols_label(ISO3 = md("**ISO3**"),
               gini_year = md("**GINI Year**"),
               gini_value = md("**GINI Value**"),
               geolocation_latitude = md("**Latitude**"),
               geolocation_longitude = md("**Longitude**"),
               religion_name = md("**Religions**"),
               religion_population = md("**Religion Populations**"),
               religion_percentage = md("**Religion Percentages**"),
               government_type = md("**Government Type**"),
               government_leaders = md("**Government Leaders**"),
               hdi = md("**Human Development Index**"),
               borders = md("**Border Nations**"),
               area = md("**Land Area**")) |>
  fmt_number(decimals = 3, use_seps = TRUE) |>
  tab_caption(md("Table 1. Variables left in lists have varying lengths"))
```

These data were pulled from the [REST Countries API](https://restcountries.com/). This is country-level data with 45 potential fields. We obtained the 10 most important variables for our analyses: country name, geolocation (longitude & latitude), population density, gdp, border nations, gini, hdi, government type, and religion. We plan on looking at measles rates related to socioeconomic variables like government, religion, hdi, gini, and gdp. We also plan on looking at measles rates related to geographic factors like border nations and geolocation. We need variables like cca3 to join to our other dataframes. Variables like border nations are maintained in lists due to their varying lengths.

# Join data

```{r}
df_year_joined <- measles_year |>
  left_join(world_df, by = c("Country Code" = "ISO3")) |>
  mutate(Density = `Total Pop.`/area) |>
  select(-ASCII)

df_year_joined |>
  head(5) |>
  gt() |> 
  tab_header(title = md("Combined Data Sets")) |>
    cols_label(Region = md("**Region**"),
               Country = md("**Country**"),
      "Country Code" = md("**ISO3**"),
      Year = md("**Year**"),
      "Total Pop." = md("**Total Pop.**"),
      "Annualized Pop. 2025" = md("**Annualized Pop. 2025**"),
      Density = md("**Population Density** ($People/{km}^2$)"),
      "Total Suspected Cases" = md("**Total Suspected Cases**"),
      "Total Measles" = md("**Total Measles Cases**"),
      "Total Lab Measles" = md("**Total Lab Measles**"),
      "Total Epi Measles" = md("**Total Epidem Measles**"),
      "Total Clinical Measles" = md("**Clinical Measles Cases**"),
      "Total Measles per Million" = md("**Measles Rate Per Million**"),
      "Total Discarded" = md("**Discarded Cases**"),
      "Total Discarded per Million" = md("**Discarded Rate Per Million**"),
               gini_year = md("**GINI Year**"),
               gini_value = md("**GINI Value**"),
               geolocation_latitude = md("**Latitude**"),
               geolocation_longitude = md("**Longitude**"),
               religion_name = md("**Religions**"),
               religion_population = md("**Religion Populations**"),
               religion_percentage = md("**Religion Percentages**"),
               government_type = md("**Government Type**"),
               government_leaders = md("**Government Leaders**"),
               hdi = md("**Human Development Index**"),
               borders = md("**Border Nations**"),
               area = md("**Land Area**")) |>
  fmt_number(decimals = 3, use_seps = TRUE) |>
  tab_caption(md("Table 2. REST API Data joined with annual measles data"))
```


# Summarize Data: Are Population Density and Measles Rate Associated?
```{r}
df_year_joined |>
  filter(`Total Pop.` >= 100000) |>
  slice_max(Density, n = 5) |>
  select(Country, Year, `Total Pop.`, Density, `Total Measles per Million`) |>
  mutate(Year = as.factor(Year)) |>
  gt() |>
  tab_header(title = md("Most Population-Dense Countries")) |>
  fmt_number(decimals = 3, use_seps = TRUE) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_caption("Table 3. Density is calculated as people per square kilometer")

df_year_joined |>
  slice_max(`Total Measles per Million`, n = 5) |>
  select(Country, Year, `Total Pop.`, Density, `Total Measles per Million`) |>
  gt() |>
  tab_header(title = md("Countries With Greatest Measles Rate")) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_caption("Table 4. Countries with high measles rates appear to have relatively low population densities")
  
```

