---
title: "project-1-checkpoint2"
author: "Cameron An"
format: html
editor: visual
code-tools: true
embed-resources: true
---

## Data

```{r}
#| message: false

# load packages
library(tidytuesdayR)
library(tidyverse)
library(stringi)
library(gt)

# load data
cases_month <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_month.csv')

cases_year <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-24/cases_year.csv')
```

## Data Cleaning

```{r}
# create measles data
measles <- cases_month |>
  select(-c(rubella_clinical, rubella_epi_linked, 
            rubella_lab_confirmed, rubella_total)) |>
  rename("Region" = region,
         "Country" = country, 
         "Country Code" = iso3,
         "Year" = year,
         "Month" = month,
         "Suspected Measles" = measles_suspect,
         "Clinical Measles" = measles_clinical,
         "Epi Measles" = measles_epi_linked,
         "Lab Measles" = measles_lab_confirmed,
         "Total Measles" = measles_total,
         "Discarded" = discarded) |>
  mutate(Date = my(str_c(Month, "-", Year)),
         ASCII = stri_trans_general(Country, "Latin-ASCII")) |>
  select(-c(Year, Month)) |>
  relocate(Date)
```

```{r}
# create measles data for years
measles_year <- cases_year |>
  select(-c(rubella_clinical, rubella_epi_linked, 
            rubella_lab_confirmed, rubella_total,
            rubella_incidence_rate_per_1000000_total_population)) |>
  rename("Region" = region,
         "Country" = country, 
         "Country Code" = iso3,
         "Year" = year,
         "Total Pop." = total_population,
         "Annualized Pop. 2025" = annualized_population_most_recent_year_only,
         "Total Suspected Cases" = total_suspected_measles_rubella_cases,
         "Total Measles" = measles_total,
         "Total Clinical Measles" = measles_clinical,
         "Total Epi Measles" = measles_epi_linked,
         "Total Lab Measles" = measles_lab_confirmed,
         "Total Measles" = measles_total,
         "Total Measles per Million" = measles_incidence_rate_per_1000000_total_population,
         "Total Discarded" = discarded_cases,
         "Total Discarded per Million" = discarded_non_measles_rubella_cases_per_100000_total_population) |>
  mutate(ASCII = stri_trans_general(Country, "Latin-ASCII")) |>
  slice(-1)
```

## Functions

### **monthly_measles_summary()**

**monthly_measles_summary():** when given the measles dataframe, country name and a given year, it outputs the total number of measles cases for each month of that year.

**Note:** the country name can be spelled changing special characters (i.e. Turkiye)

-   summary using existing columns produced with custom function

```{r}
monthly_measles_summary <- function(df, country_name, year_input) {
  
  # input checks
  required_cols <- c("Date", "Country", "Total Measles", "ASCII")
  if (!all(required_cols %in% names(df))) {
    missing_cols <- required_cols[!required_cols %in% names(df)]
    stop(paste("Missing required columns:", paste(missing_cols, 
                                                  collapse = ", ")))}
  
  if (!is.character(country_name) || length(country_name) != 1) {
    stop("Country Name must be a valid country.")}
  
  if (!is.numeric(year_input) || length(year_input) != 1) {
    stop("Year input must be a valid year.")}
  
  # empty dataset
  if (nrow(df) == 0) {
    warning("No data found.")}
  
  # handle special chars
  country_name_ascii <- stri_trans_general(country_name, "Latin-ASCII")
  
  # find the original country spelling from the dataset
  original_name <- df |>
    filter(ASCII == country_name_ascii) |>
    pull(Country) |>
    unique()
  
  # create summary table
  df |>
    mutate(Year = year(Date),
           Month = month(Date, label = TRUE, abbr = FALSE)) |>
    filter(ASCII == country_name_ascii,
           Year == year_input) |>
    group_by(Month) |>
    summarise("Total Measles Cases" = sum(`Total Measles`, na.rm = TRUE),
              .groups = "drop") |>
    arrange(Month) |>
    gt() |>
    tab_header(title = md(paste0("Monthly Measles Cases in ", original_name, 
                                " for the Year ", year_input))) |>
    cols_label(Month = md("**Month**"),
               "Total Measles Cases" = md("**Total Measles Cases**"))}
```

```{r}
# test cases (using countries with special characters)
monthly_measles_summary(measles, "Turkiye", 2012)
monthly_measles_summary(measles, "Cote d'Ivoire", 2013)
monthly_measles_summary(measles, "Chad", 2015)
```

### **yoy_change()**

**yoy_change():** when given the measles dataframe and desired country name, it outputs the annual change in measles' cases rate.

**Note:** the country name can be spelled changing special characters (i.e. Turkiye)

-   summary using modified columns produced with custom function

```{r}
yoy_change <- function(df, country_name) {
  
  # input checks
  required_cols <- c("Year", "Country", "Total Measles per Million", "ASCII")
  
  if (!all(required_cols %in% names(df))) {
    missing_cols <- required_cols[!required_cols %in% names(df)]
    stop(paste("Missing required columns:", paste(missing_cols, collapse = ", ")))}
  
  if (!is.character(country_name) || length(country_name) != 1) {
    stop("Country Name must be a valid country.")}
  
  # empty dataset
  if (nrow(df) == 0) {
    warning("No data found.")}
  
  # handle special chars
  country_name_ascii <- stri_trans_general(country_name, "Latin-ASCII")
  
  # find the original country spelling from the dataset
  original_name <- df |>
    filter(ASCII == country_name_ascii) |>
    pull(Country) |>
    unique()
  
  # create summary table
  df |>
    filter(ASCII == country_name_ascii) |>
    arrange(Year) |> 
    mutate(`Annual Change in Measles Rate` = `Total Measles per Million` - 
             lag(`Total Measles per Million`)) |>
    select(Year, `Total Pop.`, `Total Measles per Million`, 
           `Annual Change in Measles Rate`) |>
    gt() |>
    tab_header(title = md(paste0("Annual Change in Measles Cases Rate per Million in ", 
                                 original_name))) |>
    cols_label(Year = md("**Year**"),
               "Total Pop." = md("**Total Pop.**"),
               "Total Measles per Million" = md("**Total Measles per Million**"),
               "Annual Change in Measles Rate" = md("**Annual Change in Measles Rate**"))}
```

```{r}
# test cases
yoy_change(measles_year, "Chad")
yoy_change(measles_year, "Turkiye")
yoy_change(measles_year, "China")
```

### **summarize_country_seasons()**

**summarize_country_seasons():** when given the measles dataframe and desired countries, it outputs the total number of measles cases for the countries across different seasons for the entire data collection time period

**Note:** the country name can be spelled changing special characters (i.e. Turkiye)

-   summary using modified columns produced with custom function and iteration

```{r}
# southern hemisphere countries in dataset (obtained through AI)
southern_hemisphere <- c(
  "Angola", "Botswana", "Burundi", "Comoros", "Eswatini", "Lesotho", 
  "Madagascar", "Malawi", "Mozambique", "Namibia", "Rwanda", 
  "South Africa", "Zambia", "Zimbabwe", "Argentina", 
  "Bolivia (Plurinational State of)", "Brazil", "Chile", "Paraguay", 
  "Peru", "Uruguay", "Indonesia", "Timor-Leste", "Australia", 
  "Cook Islands", "Fiji", "New Zealand", "Niue", "Papua New Guinea", 
  "Samoa", "Solomon Islands", "Tonga", "Tuvalu", "Vanuatu")

# create season variable
measles_seasons <- measles |>
  mutate(Month = month(Date),
         Season = case_when(
                  # southern hemisphere
                  Country %in% southern_hemisphere & Month %in% c(12, 1, 2)  ~ "Summer",
                  Country %in% southern_hemisphere & Month %in% c(3, 4, 5)   ~ "Autumn",
                  Country %in% southern_hemisphere & Month %in% c(6, 7, 8)   ~ "Winter",
                  Country %in% southern_hemisphere & Month %in% c(9, 10, 11) ~ "Spring",
    
                  # northern hemisphere
                  Month %in% c(12, 1, 2)  ~ "Winter",
                  Month %in% c(3, 4, 5)   ~ "Spring",
                  Month %in% c(6, 7, 8)   ~ "Summer",
                  Month %in% c(9, 10, 11) ~ "Autumn")) |>
  select(-Month)
```

```{r}
summarize_country_seasons <- function(df, country_name) {
   # input checks
  required_cols <- c("Season", "Country", "Total Measles", "ASCII")
  
  if (!all(required_cols %in% names(df))) {
    missing_cols <- required_cols[!required_cols %in% names(df)]
    stop(paste("Missing required columns:", paste(missing_cols, collapse = ", ")))}
  
  if (!is.character(country_name) || length(country_name) != 1) {
    stop("Country Name must be a valid country.")}
  
  # empty dataset
  if (nrow(df) == 0) {
    warning("No data found.")}
  
  # handle special characters
  country_ascii <- stri_trans_general(country_name, "Latin-ASCII")
  
  # get proportions
  df |>
    filter(ASCII == country_ascii) |>
    group_by(Country, Season) |>
    summarise(Total_Cases = sum(`Total Measles`, na.rm = TRUE), 
              .groups = "drop") |>
    mutate(Proportion = round((Total_Cases / sum(Total_Cases)), 4)) |>
    select(-Total_Cases)}
```

```{r}
# choose countries
selected_countries <- c("Brazil", "Turkiye", "South Africa", "Australia")

# map data
map_df(selected_countries, 
       ~summarize_country_seasons(df = measles_seasons,
                                  country_name = .x)) |>
  pivot_wider(names_from = Season, 
              values_from = Proportion,
              values_fill = NA) |>
  gt() |>
  tab_header(title = md(paste0("Proportion of Measles Cases Across Different Seasons"))) |>
    cols_label(Country = md("**Country**"),
               "Autumn" = md("**Autumn**"),
               "Spring" = md("**Spring**"),
               "Summer" = md("**Summer**"),
               "Winter" = md("**Winter**")) 
```

### summ_df()

**summ_df():** when given the measles year dataframe, the function returns a dataframe summary that takes specific columns and lists the IQR, mean, median, and standard deviation of numeric variables and the unique number of levels for categorical variables.

-   summary using existing columns produced with custom function and iteration

```{r}
summ_stats <- function(x){
  if (is.numeric(x)){
    tibble(Mean = round(mean(x, na.rm = T), 4),
           Median = round(median(x, na.rm = T), 4),
           `Standard Deviation` = round(sd(x, na.rm = T), 4),
           IQR = round(IQR(x, na.rm = T), 4))}
  
  else if (is.factor(x) || is.character(x)){
    tibble(Levels = length(unique(x)))}
  
  else {
    warning("Unsupported input type")
    return(NULL)}}

summ_df <- function(df){
  if (is.data.frame(df)){
    df |>
      mutate(Year = as.factor(Year)) |>
      map(summ_stats) |>
      bind_rows(.id = "Variable") |>
      pivot_longer(cols = !Variable,
                   names_to = "Summary") |>
      pivot_wider(names_from = Variable) |>
      select(Summary, Country, Year, `Total Pop.`, 
             `Total Suspected Cases`, `Total Measles per Million`) |>
      gt() |>
      tab_header(title = md("**Data Frame Summary**")) |>
      cols_label(Summary = md("**Summary**"),
                 Country = md("**Country**"),
                 Year = md("**Year**"),
                 `Total Pop.` = md("**Total Population**"),
                 `Total Suspected Cases` = md("**Total Suspected Measles and Rubella Cases**"),
                 `Total Measles per Million` = md("**Measles Rate Per Million**"))}
  else {
    warning("Input type not supported")
    return(NULL)}}

summ_df(measles_year)
```