---
title: "project1-checkpoint4"
author: "Ruben Jimenez"
format: html
editor: visual
---

## Package loading & data import

```{r, output = FALSE, message = FALSE}
library(tidyverse)
library(rnaturalearth)
library(sf)
library(gt)
library(httr)
library(tidyjson)
library(jsonlite)
library(glue)
library(car)
library(lme4)
library(pscl)
library(fitdistrplus)
library(broom)

measles_final <- read_csv("measles_final.csv")
```

## Iterative Model Fitting

We started with a null Poisson generalized linear regression model with total measles cases as a response, as this can be modeled as count data. We aimed to maximize negative log-likelihood through the iterative addition of year-level predictor variables. While initially including population density, primary religion, nominal gdp per capita, region, year, and measles vaccine coverage as our predictors, we found that religion and region were highly multicollinear according to their variance inflation factor, meaning most of the variability in total measles cases from religion could be explained by the region. Our final model included population density, nominal gdp per capita, region, year, and MCV1 coverage, and used a negative binomial model rather than Poisson due to detected overdispersion.

```{r, warning=FALSE, output=FALSE}

model_data <- measles_final |>
  mutate(cPopDensity = scale(`Density`, scale = F),
         cNomGDP = scale(`Nominal GDP per Capita`, scale = F),
         cYear = scale(Year, scale = F),
         cMCV1 = scale(`MCV1 Coverage`, scale = F))

m0 <- glm(`Total Measles` ~ 1, family = poisson(link = "log"), data = model_data)

m1 <- glm(`Total Measles` ~ cPopDensity + `Primary Religion` + cNomGDP + Region + cYear + cMCV1, family = poisson(link = "log"), data = model_data)

m2 <- glm(`Total Measles` ~ cPopDensity + Region + cYear + cMCV1, family = poisson(link = "log"), data = model_data)

m3 <- zeroinfl(`Total Measles` ~ cPopDensity + Region + cMCV1, dist = "poisson", data = model_data)

m4 <- glm(`Total Measles` ~ cPopDensity + cNomGDP + Region + cYear + cMCV1, family = poisson(link = "log"), data = model_data)

m5 <- glm.nb(`Total Measles` ~ cPopDensity + cNomGDP + Region + cYear + cMCV1, data = model_data)
odTest(m5)

summary(m0)
summary(m1)
summary(m2)
summary(m3)
summary(m4)
summary(m5)

logLik(m1)
logLik(m2)
logLik(m3)
logLik(m4)
logLik(m5)

vif(m1)
vif(m2)
vif(m4)
```

## Final Model – Negative Binomial GLM

Based on the results of model fitting, we found that Southeast Asia has the greatest mean number of measles cases annually after adjusting for population density, nominal GDP per capita, year, and MCV1 coverage. The Americas have the lowest mean number of measles cases annually after adjusting for population density, nominal GDP per capita, year, and MCV1 coverage. Population density has a negative effect on mean measles cases after adjusting for the other variables, while nominal GDP per capita has a small negative effect on mean measles cases after adjusting for the other variables. Increases in year are associated with an increase in the mean number of measles cases after adjusting for the other variables. MCV1 coverage, or vaccination rate, has the largest negative effect on mean measles cases after adjusting for the other variables.

```{r}

tidy(m5) |>
  mutate(term = str_replace_all(term, c(
    "Region" = "",
    "cPopDensity" = "Pop. Density*",
    "cNomGDP" = "Nominal GDP per Capita*",
    "cYear" = "Year*",
    "cMCV1" = "MCV1 Coverage*",
    "\\(|\\)" = ""
  ))) |>
  rename("Model Term" = "term",
         "Estimate" = "estimate",
         "Std. Error" = "std.error",
         "Z-Statistic" = "statistic",
         "p-value" = "p.value") |>
  gt() |>
  fmt_number(decimals = 3) |>
  tab_header(
    title = md("**NB GLM Coefficients**")
  )  |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) |>
  tab_caption("Table 1: Negative Binomial regression results. Response measures total measles cases across a region in a given year. Baseline region is Africa.") |>
  tab_footnote("*Quantitative variables are centered for interpretability")
```
